---
title : "Topic Modeling of Course Content"
author: "Raphaël Morsomme"
date  : "`r Sys.Date()`"
output:
  github_document:
    toc: TRUE
---

```{r library, message = F}
library(tidyverse)
library(tidytext)

library(hunspell)    # Stemmer
library(topicmodels) # Latent Dirichlet Allocation algorithm
library(tm)          # Corpus()
```

```{r, echo = FALSE}
load("data_prep.RDATA")
```

# Introduction
University departments often have little knowledge of the actual content of their study programs. Yet, having a good understanding of what each course of a program covers is paramount to maintain and education of quality.

In this script, I conduct a topic modeling exercise on the curriculum offered by the University College Maastricht (UCM), Maastricht University, the Netherlands. UCM offers a bachelor in Liberal Arts and Science. Its curriculum contains over two hundred courses on virtually every topic conceivable - ranging from Ranging from artificial intelligence to Shakespeare and terrorism - making it a great subject for a topic modeling exercise.

The analysis is exploratory in nature: instead of answering a specific research question, I explore the data to obtain a better understanding of the content of UCM's curriculum. To accomplish this, I conduct three analyses. First, I use the tf-idf to identify the most distcinctive terms of each course. Second, I compare the curriculum across years to identify which themes have emerged and declined. Finally, I use the Latent Dirichlet Allocation algorithm to create a topic model of the 2018-2019 curriculum, both at the course- and at the cluster-level. The textual data are the course descriptions present in the course catalogues.

# Data Preparation
## Overview
We starting by importing the course catalogues and a data set containing information at the course level. We then extract the descriptions of the courses from the course catalogues. These descriptions are one to two pages long and form the textual data that we will analyze. Lastly, we transform the data into the *[tidy text format](https://www.tidytextmining.com/tidytext.html)*, stem the terms and remove the stop-words.

## Importing
We import two datasets from the directory. `d_course` is a tiblle indicating the code, name and cluster of each course. (Courses are distributed among 17 clusters: International Relation, Cultural Studies, Biomedical Science, etc.). `corpus` is a corpus containing the five most recent course catalogues of UCM. Course catalogues are published every year and contain a descritpion of each course that is one or two pages long.

```{r}
d_course <- read_csv("Course.csv", col_type = cols())
```

```{r import data, eval = FALSE}
corpus <- Corpus(x             = DirSource("Catalogues"),
                 readerControl = list(reader = readPDF(control = list(text = "-layout"))))
```

## Extracting Course Descriptions from Course Catalogues
We extract the description of each course from the course catalogues. The code is a little longish and does not add much to the script, so I included it in a seperate appendix.

```{r description, eval = FALSE, echo = FALSE}
# Setup
n_catalogue    <- length(corpus)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description  <- tibble(Code            = character(0),
                         `Calendar Year` = character(0),
                         Description     = character(0))

course_code <- c("COR", "HUM", "SCI", "SSC", "SKI", "PRO", "UGR", "CAP")
course_code <- paste(course_code, collapse = "|")

content_to_exclude <- c(
  "^Core Courses \\(COR\\)",
  "^Humanities \\(HUM\\)",
  "^Sciences \\(SCI\\)",
  "^Social Sciences \\(SSC\\)",
  "^Skills Trainings \\(SKI\\)", "^Skills Training \\(SKI\\)",
  "^Project \\(PRO\\)", "^Projects \\(PRO\\)",
  "^Undergraduate Research \\(UGR\\)",
  " UCM Undergraduate\r\nResearch", "UCM Undergraduate\r\n  Research"
  )

content_to_exclude <- paste(content_to_exclude, collapse = "|")

# Loop 1
for(n in 1 : n_catalogue){
  
  cat <- content(corpus[[n]])
  
  # Description Section
  page_description_start <- grep(pattern = "^Core Courses \\(COR\\)", x = cat) + 1
  page_description_end   <- grep(pattern = "Appendix"               , x = cat) - 1
    if(length(page_description_end) == 0) page_description_end <- length(cat)
  cat_description <- cat[page_description_start : page_description_end]
  
  # Course Descriptions Only
  pages_to_exclude <- grep(pattern = content_to_exclude, x = cat_description)
  cat_description  <- cat_description[!1:length(cat_description) %in% pages_to_exclude]
  
  # First Pages of Course Descriptions
  first_three_letters     <- substr(cat_description, start = 1, stop = 3)
  first_pages_description <- grep(pattern = course_code, x = first_three_letters)

  # Loop 2
  for(page in first_pages_description){
    
    # Extract Description
    following_page <- page + 1 # for convenience
    description <- if(following_page %in% first_pages_description) cat_description[page]
                   else                                            paste(cat_description[page : following_page], collapse = " ")
    
    # Save Description
    Code <- substring(description, first = 1, last = 7)      
    year <- calendar_years[n]
    d_description <- d_description %>%
      add_row(Code            = Code,
              `Calendar Year` = year,
              Description     = description)
    
  } # close for-loop 2 (page)
} # close for-loop 1 (n)
```

```{r}
print(d_description)
```

## Tidying
We save the course descriptions in the [tidy text format ](https://www.tidytextmining.com/tidytext.html) with one row per course-year-term.
```{r tidy, eval = FALSE}
d_description_tidy <- unnest_tokens(d_description, output = word, input = Description)
```

```{r}
print(d_description_tidy)
```

## Stemming
Lastly, we stem the terms and filter out stop words. We use the stemmer from the `hunspell` package to build a stemming function `stem_hunspell()` which takes a term as input and returns its stem. We prefer the Hunspell stemmer over the usual Snowball stemmer because it offers a more precise stemming.

> **Trick: dictionary-based approach to stem a large number of terms.**

> Since it would take too much time to apply our stemming function to all `340.000` terms of `d_description_tidy`, we use a *dictionary-based approach*. We create a dictionnary that provides the stem of the `8,500` unique terms present in the dataset and then `jull_join` the newly created dictionary and `d_description_tidy` to stem all the terms at once. This way, we greatly reduce the number of times we use the stemming function.

```{r stem, eval = FALSE}

# Stemming function
stem_hunspell <- function(term) {
  # look up the term in the dictionary
  stems <- hunspell_stem(term)[[1]]
  
  # identify the stem
  if (length(stems) == 0) { # if no stem in dictionary, use  original term
    stem <- term
  } else { # if multiple stems, use last one (most basic)
    stem <- stems[[length(stems)]]
  }
  
}

# Dictionary 
my_dictionary <- d_description_tidy %>%
  distinct(word) %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))

# Full join
d_description_stem <- d_description_tidy %>%
  full_join(my_dictionary, by = "word") %>%
  rename(word_original = word,
         word          = word_stem) %>%
  # Removing stop words
  filter(!word %in% stop_words$word,
         !word %in% as.character(1:1e3))
```

```{r}
print(d_description_stem) # See humanities (original) - humanity (stem)
```

```{r, echo =  FALSE, eval = FALSE}
save(d_description, d_description_tidy, d_description_stem, file = "data_prep.RDATA")
```

# Analysis

## Overview
We conduct three analyses. First, we identify the most important terms of each course and cluster with the tf-idf. Next, compare the curriculum across years to identify which themes have emerged and declined. Finally, we use the LDA algorithm (a popular technique for topic modeling) to build a topic model of the 2018-2019 curriculum, both at the course- and cluster-level.

## Most Important Terms
The tf-idf is a popular measure to identify the most important terms of documents that belong to a corpus. By penalizing terms that occur in many documents, tf-idf allows us to focus on the terms that are specific to each document. In our case, terms such as "learn" or "student", which appear in a large number of course descriptions and are therefore not informative, have a low tf-idf. This way, we can identify the most *distinctive* terms of each course/cluster and get a feel of the topics they cover.

We use the function `bind_tf_idf()` to obtain the tf-idf of each term for the year 2018-2019. We then identify the most important terms of each cluster and course[^9] and display them both as barplots and word clouds. In the latter, the size and the color of a term indicates its tf-idf.

```{r tdm}
tf_idf <- d_description_stem %>%
  filter(`Calendar Year` == "2018-2019") %>%
  count(Code, word, sort = T) %>%
  bind_tf_idf(term = word, document = Code, n = n) %>%
  left_join(d_course, by = "Code")
  

print(tf_idf)

tf_idf_cluster <- tf_idf %>%
  group_by(Cluster, word) %>%
  summarise(tf_idf = sum(tf_idf),
            n      = sum(n)) %>%
  ungroup
```

### Results
The tf-idf does a pretty good job at identifying the most important terms of the courses/clusters. If the names of the courses/clusters were absent from the plots, it would be fairly easy to guess them from the terms. We also observe interesting elements concerning the content of the clusters. For instance, while the cluster `History` gives a central place to the European continent (with terms like Europe, european), the cluster `International Relations` focuses more on China (chinese).

## Term Emergence
We compare the content of the curriculum across years to identify terms and themes that have declined and emerged. To accomplish this, we compare the frequency of the terms in the two catalogues by taking their `log ratio`. A positive value indicates that the term has become more frequent and a negative value indicates a decline in the use of the term. We plot the forty terms with the highest absolute `log ratio`. Again, we display the information both as a barplot and wordcloud.

### Results

Let us compare the years `2014-2015` (oldest available) and `2018-2019` (most recent). From the two plots, we can observe that topics that have made the news these last few years such as religion, migrationa and climate have become more important in the catalogues. This shows that the college has done a good job at adapting its curriculum to the development of the world's affairs. We also observe that the themes of sexuality (with terms like `sex`, `sexual` and `sexuality`) and terrorism (`terrorism`, `crime`) have declined. It is interesting to note that the terms `globalization`, `poverty` and `sustainability` have become less important. As for the term `portal`, its "emergence" is due to the introduction of a new online student *portal* system at the college. The terms is not mentioned once in 2014-2015 and appears `19` times in 2018-2019. This shows that human interpretation and a good knowledge of the data is crucial in such analysis to avoid false alarms.

## Topic Modeling
Finally, we use the Latent Dirichlet Allocation (LDA) algorithm to conduct a topic analysis of the college's curriculum at the course- and the cluster-level. Given a corpus of documents and a predetermined number of topics, the LDA algorithm outputs a topic model which gives the importance of each term to the topics (beta distribution) and the importance of each topic to the documents (gamma distribution). In other words, the LDA finds the mixture of words associated with each topic and the mixture of topics associated with each document. The advantage of the LDA over regular clustering methods is that is allows for overlap of terms across topics and of topics across documents, thereby offering a model that is closer to natural language.

### Fitting Model
We build four models with respectively `5`, `12`, `17` and `25` topics for the year `2018-2019`.

TODO: use package `ldatuning` to determine best number of topics. (https://cran.r-project.org/web/packages/ldatuning/vignettes/topics.html)

```{r cast, eval = F}
d_cast <- d_description_stem %>%
  filter(`Calendar Year` == "2018-2019") %>%
  count(Code, word) %>%
  cast_dtm(Code,word, n)
```

```{r, fitting LDA, eval = FALSE}
LDA_5  <- LDA(d_cast, k = 5 , control = list(seed = 123))
LDA_12 <- LDA(d_cast, k = 12, control = list(seed = 123))
LDA_17 <- LDA(d_cast, k = 17, control = list(seed = 123)) # There are 17 clusters
LDA_25 <- LDA(d_cast, k = 25, control = list(seed = 123))
```

### Results
We present the output of a topic model with three plots which respectively show (i) the most important terms for each topic (ii) the main courses/clusters of each topic and (iii) the most important topics of each course/cluster. For each model, we present the results at the course- an the cluster-level. The LDA is an unsupervised algorithm, meaning that the topic are unlabeled. In the final section, I label the topic manually.

Let us set the number of topics to `12` and observe the plots at the cluster level. Most topics are covered by several clusters. Fot instance, `topic 10` is covered in several clusters of the humanities (`History`, `Literature` and `Philosophy`) and social sciences (`International Relations`, `International Law` and `Economics`). For a liberal arts program this is a desirable outcome since it encourages students interested in a particular topic to take classes in different clusters, thereby broading their acadmeic horizon (one of the objectives of the program). At the same time, this pattern may be artificially created by the fact that there are more than twelve topics present in the curriculum, meaning that the LDA algorithm has to combine unrelated themes into one topic. This is for instance the case for `topic 3` which combines the themes of law (`law`, `legal`) and international affairs (`european`, `international`) (see first plot). Increasing the number of topics in the model should solve this issue.

We also observe that the distribution of topics is very different at the course and the cluster level. While courses are usually heavily dominated by a single topics, clusters contain several major topics. Looking at the second plot, we observe that the courses `Computer Science`, `Optimization` and `Philosophy of Language` for instance are heavily dominated by `topic 7`. The fourth plots shows that most clusters contain several topics. Interestingly, this graph shows that the same topic (`topic 10`) dominates the clusters `History`, `International Relations` and `Literature`. This indicates either that the content of the three clusters share some similarity or that `topic 10` contains several different themes.

So far, we have only analyzed models with unlabelled topics. In order to give more substance to our analysis, we assign labels to the topics. To accomplish this, we look at the first plot showing the most important terms of each topic and attempt to identify the common theme(s) among them. Some topics are easier to label than others. `topic 6` is for instance pretty straightforward: it is dominated by the term `search` and also contain the terms `qualitative`, `read`, `interview` and `study`: `topic 6` corresponds to qualitative research skills and we therefore label it `Qual. Res.`. Labelling `topic 7` on the other hand is more trick. The best label I could find is `Engineering`. I have labelled each topic and present the results under the tab `labeled topics`.

The labels give us a better image of the actual content of the courses and the clusters. The last plot indicate thta most clusters cover the topics that we expect them to cover, indicating that the current division of courses in clusters is backed up by the content of the courses. As expected, the cluster `Sociology` covers the topics of `Society` and `Culture` and the cluster `Cultural Studies` covers the topics of `Arts`, `Culture` and `Qual. Res.`. Yet, I am surprised by the absence of certain topics in some clusters. For instance, the cluster `history` lacks the topics of `culture` and `society` and the topic of `research` is also barely present in `Biomedical Sciences`. The former shows the heavy focus of the `history` cluster on war and conflicts (`Foreign Policy`) and the latter reflects the absence of research projects in the classes of the biomedical cluster, something the college might want to investigate.

```{r, eval = FALSE, echo = FALSE}
LDA_5 = list(beta = tidy(LDA_5, matrix = "beta"),
             gamma = tidy(LDA_5, matrix = "gamma"))
LDA_12 = list(beta = tidy(LDA_12, matrix = "beta"),
             gamma = tidy(LDA_12, matrix = "gamma"))
LDA_17 = list(beta = tidy(LDA_17, matrix = "beta"),
             gamma = tidy(LDA_17, matrix = "gamma"))
LDA_25 = list(beta = tidy(LDA_25, matrix = "beta"),
             gamma = tidy(LDA_25, matrix = "gamma"))

save(
  
  # TF-IDF
  tf_idf, tf_idf_cluster,
  
  # Topic emergence
  d_description_stem,
  
  # Topic nodeling
  d_course, LDA_5, LDA_12, LDA_17, LDA_25,
  
  # File name
  file = "Shiny App - Topic Modeling/results_data.RDATA"
  )

```
